---
title: "Regression Demo"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(knitr)
library(ggplot2)
library(mnormt)
library(LearnBayes)
library(coda)
library(rjags)
library(yarrr)
```

## Data Description
The data was obtained from [https://www.kaggle.com/fivethirtyeight/the-ultimate-halloween-candy-power-ranking](https://www.kaggle.com/fivethirtyeight/the-ultimate-halloween-candy-power-ranking). The original article and analysis is available at [https://fivethirtyeight.com/features/the-ultimate-halloween-candy-power-ranking/](https://fivethirtyeight.com/features/the-ultimate-halloween-candy-power-ranking/).

```{r}
candy <- read_csv('http://math.montana.edu/ahoegh/teaching/stat446/candy-data.csv')
candy <- candy %>% mutate(chocolate_factor = as.factor(chocolate), 
                          nut_factor = as.factor(peanutyalmondy))
```

##### Context

What’s the best (or at least the most popular) Halloween candy? That was the question this dataset was collected to answer. Data was collected by creating a website where participants were shown presenting two fun-sized candies and asked to click on the one they would prefer to receive. In total, more than 269 thousand votes were collected from 8,371 different IP addresses.

##### Content

candy-data.csv includes attributes for each candy along with its ranking. For binary variables, 1 means yes, 0 means no. The data contains the following fields:

- chocolate: Does it contain chocolate?
- fruity: Is it fruit flavored?
- caramel: Is there caramel in the candy?
- peanutalmondy: Does it contain peanuts, peanut butter or almonds?
- nougat: Does it contain nougat?
- crispedricewafer: Does it contain crisped rice, wafers, or a cookie component?
- hard: Is it a hard candy?
- bar: Is it a candy bar?
- pluribus: Is it one of many candies in a bag or box?
- sugarpercent: The percentile of sugar it falls under within the data set.
- pricepercent: The unit price percentile compared to the rest of the set.
- winpercent: The overall win percentage according to 269,000 matchups.

##### Acknowledgements:

This dataset is Copyright (c) 2014 ESPN Internet Ventures and distributed under an MIT license. Check out the analysis and write-up here: The Ultimate Halloween Candy Power Ranking. Thanks to Walt Hickey for making the data available.


## Data Exploration

```{r}
candy %>% ggplot(aes(x = winpercent)) + geom_histogram() + ggtitle('Distribution for Win Percent') + 
  xlim(0,100)
candy %>% ggplot(aes(y = winpercent, x = chocolate_factor)) + geom_violin() + geom_jitter() + 
  xlab('Chocolate') + ylim(0,100)
candy %>% ggplot(aes(y = winpercent, x = nut_factor)) + geom_violin() + geom_jitter() + 
  xlab('Peanut or Almond') + ylim(0,100)
```


## Linear Models Demo

The goal of this analysis will be to model the `winpercent` variables in the dataset. I'll use the `chocolate` and `peanutyalmondy` variables, but you are welcome to try your own covariates.

### Linear Model Framework

###### Sampling Model
$$\tilde{Y} \sim N(X \tilde{\beta}, \sigma^2 I)$$
where $\tilde{Y}$ is a vector of length n = `r nrow(candy)` where the $i^{th}$ response is the winpercent of candy $i$,   $$X = \begin{bmatrix}
  1&chocolate_1& peanutyalmondy_1 \\
  1&chocolate_2 &peanutyalmondy_2 \\
  \vdots &  \ddots & \vdots\\
  1&chocolate_n &peanutyalmondy_n
  \end{bmatrix}$$

###### Priors

$$\tilde{\beta}\sim N(\tilde{\beta_0}, \Sigma_0)$$
$$\sigma^2 \sim IG \left(\frac{\nu_0}{2},\frac{\nu_0 \sigma^2_0} {2} \right)$$


#### 0. `lm()`

```{r}
lm_candy <- lm(winpercent ~ chocolate_factor + nut_factor, data = candy)
summary(lm_candy)
predict(lm_candy, data.frame(chocolate_factor = as.factor(c(1,1,0,0)), nut_factor = as.factor(c(1,0,1,0))))
```

- **Interpret the coefficients in this model**

#### 1. Gibbs Sampler

```{r, eval = T}
set.seed(10262018)
y <- candy$winpercent
X <- model.matrix(winpercent ~ chocolate_factor + nut_factor, data = candy)
p <- ncol(X)
n <- nrow(X)

# Initialization and Prior
num_mcmc <- 5000
beta_0 <- rep(0,p)
Sigma_0 <- diag(p) * 1000
Sigma_0_inv <- solve(Sigma_0)
nu_0 <- .02
sigmasq_0 <- 1
beta_samples <- matrix(0, nrow = num_mcmc, ncol = p)
sigmasq_samples <- rep(1, num_mcmc)

for (iter in 2:num_mcmc){
  # sample beta
  cov_beta <- solve(Sigma_0_inv + t(X) %*% X / sigmasq_samples[iter - 1])
  exp_beta <- cov_beta %*% (Sigma_0_inv %*% beta_0 + t(X) %*% y / sigmasq_samples[iter-1])
  beta_samples[iter,] <- rmnorm(1, exp_beta, cov_beta) 
  
  # sample sigmasq
  sigmasq_samples[iter] <- rigamma(1, .5 * (nu_0 + n) , 
          .5 * (nu_0 * sigmasq_0 + t(y - X %*% beta_samples[iter,]) %*% 
                  (y - X %*% beta_samples[iter,])) )
}

burn_in <- 100
beta_samples[(burn_in+1):num_mcmc,] %>% as.mcmc() %>% summary()
beta_samples[(burn_in+1):num_mcmc,] %>% as.mcmc() %>% HPDinterval()
beta_samples[(burn_in+1):num_mcmc,] %>% as.mcmc() %>% effectiveSize()

sqrt(sigmasq_samples[(burn_in+1):num_mcmc]) %>% as.mcmc() %>% summary()

tibble(beta0 = beta_samples[(burn_in+1):num_mcmc,1], iteration = 1:(num_mcmc - burn_in)) %>% 
  ggplot(aes(y = beta0,iteration)) + geom_line()

```


#### 2. JAGS


###### Model Specification
```{r}
model_string <- "model{
  # Likelihood
  for(i in 1:n){
    y[i]   ~ dnorm(mu[i],inv.var)
    mu[i] <- beta[1] + beta[2]*chocolate[i] + beta[3]*peanut[i]
  }

  
  # Note priors are hard-coded, but could be variables
  
  # Prior for beta 
  for(j in 1:3){
    beta[j] ~ dnorm(0,0.001)
  }

  # Prior for the inverse variance
  inv.var   ~ dgamma(0.01, 0.01)
  sigma     <- 1/sqrt(inv.var)

}"
```

###### Compile in JAGS
```{r}
model <- jags.model(textConnection(model_string), 
                    data = list(y = candy$winpercent,n = nrow(candy),
                                chocolate = candy$chocolate,
                                peanut=candy$peanutyalmondy))
```

###### Draw Samples

```{r}
# Burnin for 1000 samples
update(model, 1000, progress.bar="none") 

samp <- coda.samples(model, 
        variable.names=c("beta","sigma"), 
        n.iter=5000, progress.bar="none")

summary(samp)
#plot(samp)
```

#### 3. Stan

##### Model Statement

```{stan output.var="lm_stan"}
data {
  int<lower=0> N;
  vector[N] chocolate;
  vector[N] peanut;
  vector[N] y;
}
parameters {
  real beta0;
  real beta1;
  real beta2;
  real<lower=0> sigma;
}
model {
  y ~ normal(beta0 + beta1 * chocolate + beta2 * peanut, sigma);
}
```

```{r}
library(rstan)
post <- rstan::sampling(lm_stan, 
             data = list(y = candy$winpercent, N = nrow(candy), peanut = candy$peanutyalmondy, chocolate = candy$chocolate))
post
```


### Posterior Predictive Distribution
A quote from the Bayesian Data Analysis textbook highlights the goal of model checking.
>> We do not like to ask: 'Is our model true or false?', since probability models in most data analyses will not be perfectly true... The more relevant question is 'Do the model's deficiencies have a noticeable effect on the substantive inference?'

```{r, warning = F}
Y <- rep(0, (num_mcmc - burn_in))
for (i in 1:(num_mcmc - burn_in)){
  Y[i] <- rnorm(1,  beta_samples[i+burn_in,] %*% X[sample(n,1),], sqrt(sigmasq_samples[i+burn_in]))
}

tibble(Y = Y) %>% ggplot(aes(x = Y)) + geom_histogram(bins = 100) + ggtitle('Posterior Predictive Distribution') + 
  xlim(0,100)

candy %>% ggplot(aes(x = winpercent)) + geom_histogram() + ggtitle('Distribution for Win Percent') + 
  xlim(0,100)
```

